---
title: "DATA607 Final Project"
author: "Dhanya Nair"
date: "2024-04-28"
output:
  html_document: default
  pdf_document: default
  toc: true
---

## Topic for analysis:

### Tesla has reduced the price of its Electric Vehicle(EV) models 3 times within a year .

### The question we are seeking to answer is why is Tesla's EV sales and price declining though Sales were projected to grow by 30% in 2025. 
### Possible reasons can be growing competitors, more incentives, more inventory, insufficient charging infrastructure. 
### Benefit of doing this work is understanding the market conditions to make better decisions from the consumer side as well as from the auto industry side. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries:
```{r,include=FALSE, warn.conflicts = FALSE}
library(tidyverse)
library(pdftools)
```
##  Data sources:
### 1. CSV Format
### Global Electric Vehicle Dataset ( data till 2022, projections for 2025, 2030)
### File location: https://www.kaggle.com/datasets/padmapiyush/global-electric-vehicle-dataset-2023/data
```{r}


#Load the original .csv file from Global Electric Vehicle Dataset 2023
#file location: https://www.kaggle.com/datasets/padmapiyush/global-electric-vehicle-dataset-2023/data

EV_Raw <- read.csv("https://raw.githubusercontent.com/datanerddhanya/DATA607/main/finalproject/Global%20EV%202023.csv", header = TRUE)

#filter to required parameters and mode as we do not need vans, buses and trucks
EV_Data <- EV_Raw %>%
  filter( parameter == "EV sales" |parameter == "EV stock" | parameter == "EV charging points") %>%
  filter( mode == "Cars" | mode == "EV") 

# for the future years 2025 and 2030, change the category to projection
EV_Data <- EV_Data %>%
  mutate( category = ifelse ( year =="2025" | year =="2030","Projection",category  ))

# to remove the scientific display format
options(scipen = 999)

# filter the required records as we do not need projected records for the previous years
# filter to years 2021 onwards
# remove groupings within the region variable so that we retain only individual countries.
EV_Data <- EV_Data %>%
filter( category == "Projection" | category == "Historical" ) %>%
filter( year > 2020) %>%
subset(!(region %in% c('EU27', 'Europe', 'Other Europe', 'Rest of the world', 'World'))) %>%
  mutate( value = round(as.numeric(value))) %>%
  mutate( year = as.factor(as.numeric(year)))

# to find the top countries of interest
# i see that the following are the top countries in EV sales 
#"China" , "USA", "India", "Germany", "United Kingdom"
EV_Data |>
  filter( parameter == "EV sales" ) |>
  filter( mode == "Cars") |>
  group_by(region) |>
  arrange(desc(value)) |>
  ggplot( aes(x= value,y = reorder(region,value))) +
  geom_bar(stat = "identity", position = "dodge", fill = "green") +
  labs(title = "Top Countries in EV Sales for all years from 2021 till date", x= "EV Sales" , y = "Countries")
  

#hybrid vs battery vehicles
EV_Data |>
  filter( parameter == "EV sales" ) |>
  filter( mode == "Cars") |>
  ggplot( aes(x= year ,y = value , color = powertrain)) +
  geom_point(stat = "identity") +
  labs(title = "Battery Electric Vehicle vs Plugin Hybrid  Electric Vehicle", x= "Year" , y = "EV Sales")
    

# Bar plot
EV_Data |>
 filter( region %in% c("China" , "USA", "India", "Germany", "United Kingdom")) |>
ggplot( aes(x= year,y = value, fill = parameter)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ region, scales = "free") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Set3") +
  scale_x_discrete(breaks = unique(EV_Data$year)) +
labs(title = "EV Production units, Sales Units and Charging points by region by year", x= "Year" , y = "Units")


# Line plot
EV_Data |>
 filter( region %in% c("China" , "USA", "India", "Germany", "United Kingdom")) |>
ggplot(aes(x = year, y = value, color = region, group = region)) +
  geom_line() +
  facet_wrap(~ parameter, scales = "free_y") +
  labs(title = "Electric Vehicle Data Trends by Region",
       x = "Year",
       y = "Value") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Set2")

  
# Faceted plot
EV_Data |>
 filter( region %in% c("China" , "USA", "India", "Germany", "United Kingdom")) |>
ggplot(aes(x = year, y = value)) +
  geom_line() +
  facet_grid(parameter ~ region, scales = "free") +
  labs(title = "Electric Vehicle Data Trends by Parameter and Region",
       x = "Year",
       y = "Value") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Set2")

```

### Conclusion:

- Popular countries to sell EV vehicles are China, India, USA, United Kingdom .

- BEV powertrain/ Battery operated electric vehicle is becoming more and more popular every year.

- USA is projected to have huge stock of EV cars, however the EV public charging points are  relatively less. This may be a concern for consumers who prefer to have easily accessible charging points.

- Based on the bar plot, i see that the EV stock/ production is much more than the EV sales for each country. This may cause the Sale price to decrease due to unsold inventory. 


### 2. PDF Format
### Gather Tesla Production and sales units for 2023 and 2024
### File location: https://digitalassets.tesla.com/tesla-contents/image/upload/IR/TSLA-Q1-2024-Update.pdf

```{r}
#read data from the pdf file
pdf_text <- pdf_text("https://digitalassets.tesla.com/tesla-contents/image/upload/IR/TSLA-Q1-2024-Update.pdf")

# Extract text from the 6th page
page_text <- pdf_text("https://digitalassets.tesla.com/tesla-contents/image/upload/IR/TSLA-Q1-2024-Update.pdf")[6]


# Split the text into lines
lines <- strsplit(page_text, "\n")[[1]]


# Identify and extract the lines corresponding to the table
# Adjust these indices based on where the table is located in the text
table_lines <- lines[3:11]


# Assuming the table is tabular, i can then further process the lines to create a data frame
# Define a regular expression for splitting by one or more spaces
split_regex <- "\\s+"
# Split each line by the regular expression
table_data <- lapply(table_lines, function(line) strsplit(line, split_regex )[[1]])

# Convert the list to a data frame
df <- as.data.frame(do.call(rbind, table_data))

# filter only required observations
# rename columns
# concatenate two variables into one
# remove the unwanted columns
Tesla_data <- df |>
  filter(V2 == "Total") |>
  rename ( type = "V3", Q1_2023 = "V4" , Q2_2023 = "V5", Q3_2023 = "V6", Q4_2023 = "V7", Q1_2024 ="V8", YoY = "V9")  |>
  mutate ( type = paste(V2, " ", type))|>
  subset(select= -c(V1,V2,V10,YoY) )

Tesla_stats <- Tesla_data |>
  pivot_longer(
    cols = starts_with("Q"),
    names_to = "Time", 
    values_to = "Units",
   values_drop_na = TRUE
  )


#visualise
ggplot(Tesla_stats , aes(x= Time,y = Units, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
    theme_minimal() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Set3") +
  labs( title = "Tesla EV models - Production Vs Sales")



```

### Conclusion
- As we can see in the bar chart there is drop in production and sales of Tesla cars in Q1 2024 compared to Q1 2023. 
- This may have cause the price to reduce from Q3 2023, which caused a bit of sales increase in Q4 2023.

### 3. PDF Format
### Competitor Analysis
### File location: https://www.coxautoinc.com/wp-content/uploads/2024/04/Q1-2024-Kelley-Blue-Book-Electric-Vehicle-Sales-Report.pdf

```{r}
#read data from the pdf file
comp_pdf_text <- pdf_text("https://www.coxautoinc.com/wp-content/uploads/2024/04/Q1-2024-Kelley-Blue-Book-Electric-Vehicle-Sales-Report.pdf")

# Extract text from the 6th page
comp_page_text <- pdf_text("https://www.coxautoinc.com/wp-content/uploads/2024/04/Q1-2024-Kelley-Blue-Book-Electric-Vehicle-Sales-Report.pdf")[2]


# Split the text into lines
comp_lines <- strsplit(comp_page_text, "\n")[[1]]

print(comp_lines)
# Identify and extract the lines corresponding to the table
# Adjust these indices based on where the table is located in the text
comp_table_lines <- comp_lines[4:31]


# Assuming the table is tabular, i can then further process the lines to create a data frame
# Define a regular expression for splitting by one or more spaces
split_regex <- "\\s+"
# Split each line by the regular expression
comp_table_data <- lapply(comp_table_lines, function(line) strsplit(line, split_regex )[[1]])

# Convert the list to a data frame
comp_ev_sales <- as.data.frame(do.call(rbind, comp_table_data))


# filter only required observations
# rename columns
# concatenate two variables into one
# remove the unwanted columns
ev_brand_sales <- comp_ev_sales|>
  rename ( Brand = "V1", Y_2024 = "V2", Y_2023 = "V3", YOY = "V4" , Segment_share = "V5") |>
  subset(select= -c(V6) ) 

#Tidying up data in the variables to remove thousand separator, values with -
ev_brand_sales$Y_2024 <- gsub(",", "", ev_brand_sales$Y_2024)
ev_brand_sales$Y_2024 <- gsub("-", "0", ev_brand_sales$Y_2024)
ev_brand_sales$Y_2023 <- gsub(",", "", ev_brand_sales$Y_2023)
ev_brand_sales$Y_2023 <- gsub("-", "0", ev_brand_sales$Y_2023)
ev_brand_sales$Segment_share <- gsub("%", "", ev_brand_sales$Segment_share)
ev_brand_sales$YOY <- gsub("%", "", ev_brand_sales$YOY)

ev_brand_YOY <- ev_brand_sales |>
  mutate( Segment_share= as.numeric(Segment_share),Y_2024 = as.numeric(Y_2024), Y_2023 = as.numeric(Y_2023),YOY = as.numeric(YOY) ) |>
  drop_na(YOY)


# Convert Brand to a factor and reorder based on YOY %
# this is needed to sort the yoy% in order
ev_brand_YOY$Brand <- factor(ev_brand_YOY$Brand, levels = ev_brand_YOY$Brand[order(-ev_brand_YOY$YOY)])


#visualize year of year change
  ggplot(ev_brand_YOY , aes(x= Brand,y = YOY)) +
  geom_bar(stat = "identity", position = "dodge", fill = "blue") +
  coord_flip() +
    labs ( title = "Year over Year change in Sales of EV cars for 2023 and 2024" , 
           y = "YoY %" , x = "Brand")

# Convert Brand to a factor and reorder based on Segment_share %
# this is needed to sort the Segment share % in order
ev_brand_YOY$Brand <- factor(ev_brand_YOY$Brand, levels = ev_brand_YOY$Brand[order(-ev_brand_YOY$Segment_share)])

  
#visualize segment share YTD
  ggplot(ev_brand_YOY , aes(x= Brand,y = Segment_share)) +
  geom_bar(stat = "identity", position = "dodge", fill = "blue") +
  coord_flip() +
    labs ( title = "Segment share of brands in Sales of EV cars for 2023 and 2024" , 
           y = "Segment Share in %" , x = "Brand")
```

### Conclusion:
- Based on the visualized plots,we can infer from the bar plot that the competition for the EV cars has significantly increased. Many brands have significantly invested in EV cars. Tesla YoY % is -13.3 % .This may have caused Tesla to reduce the sale price of its EV cars to compete with the growing competition. 
- However, Tesla is still a leader in the EV segment with respect to overall sales.


